---
title: "Monthly Deaths from Lung Diseases in the UK"
author: "Ihar Kukharchuk"
date: "June 9, 2018"
output:
    ioslides_presentation:
        keep_md: true
---

```{r setup, include=FALSE}
library(plotly)
library(xts)
```

## Overview

Performation of exploratory analysis for three time series giving the monthly
deaths from bronchitis, emphysema and asthma in the UK, 1974–1979, both sexes
(ldeaths), males (mdeaths) and females (fdeaths).

The considering dataset is available in [R datasets][1] package with the
following [description][2].

[1]: http://stat.ethz.ch/R-manual/R-devel/library/datasets/html/00Index.html
[2]: http://stat.ethz.ch/R-manual/R-devel/library/datasets/html/UKLungDeaths.html

## Brief Exploratory Analysis

```{r}
ldeaths
summary(ldeaths)
```

## Deaths Over the Years - Code Part 1

```{r}
death.df <- data.frame(date = index(ldeaths),
                       year = round(index(ldeaths)),
                       month = c(1:12),
                       all = as.numeric(ldeaths),
                       male = as.numeric(mdeaths),
                       female = as.numeric(fdeaths))
date <- paste0("Date: ", month.abb[death.df$month],
               ", ", death.df$year, "<br>")
death.ht_all <- paste0(date, "Count: ", death.df$all)
death.ht_male <- paste0(date, "Count: ", death.df$male)
death.ht_female <- paste0(date, "Count: ", death.df$female)
```

## Deaths Over the Years - Code Part 2

```{r}
death.plot <- plot_ly(death.df, x = ~date) %>%
        add_lines(y = ~all,
                  line = list(color = "#ff0000", width = 1),
                  hoverinfo = "text", text = death.ht_all,
                  name = "All") %>%
        add_lines(y = ~male,
                  line = list(color = "#00ff00", width = 1),
                  hoverinfo = "text", text = death.ht_male,
                  name = "Male") %>%
        add_lines(y = ~female,
                  line = list(color = "#0000ff", width = 1),
                  hoverinfo = "text", text = death.ht_female,
                  name = "Female") %>%
        layout(xaxis = list(title = "timeline"),
               yaxis = list(title = "count"))
```

## Deaths Over the Years - Graph

```{r, echo = FALSE}
death.plot
```

## Deaths Summary Per Month - Code

```{r}
df <- tapply(death.df$all, death.df$month, FUN = sum)
df <- data.frame(index = c(1:12), all = df)
death.monthplot <- plot_ly(df, x = ~month.abb[index], y = ~all) %>%
        layout(xaxis = list(categoryorder = "array",
                            categoryarray = df$index,
                            title = "month"),
               yaxis = list(title = "count"))
```

## Deaths Summary Per Month - Graph

```{r, echo = FALSE, message = FALSE, warning = FALSE}
death.monthplot
```

## Summary

- the amount of deaths from lung deseases in period 1974-1980 almost the same
- the biggest amount of deaths happened during first three months of each year

## References

* [Monthly Deaths from Lung Diseases in the UK (UKLungDeathshttp dataset)][3]

[3]: http://stat.ethz.ch/R-manual/R-devel/library/datasets/html/UKLungDeaths.html